Mapping Cancer: the Potential of Cartograms and Alternative Map Displays

### 2020-05-14 literature.Rmd
begin letter to editors using reviewer feedback
New figure 1 (include ACA)

hopeful

Table 1 and 2 switched, stats updated

---
title: "Mapping Cancer: the Potential of Cartograms and Alternative Map Displays"
author: "Stephanie Kobakian^[Science and Engineering Faculty, Queensland University of Technology.], Dianne Cook^[Faculty of Econometrics and Business Statistics, Monash University.] and Jessie Roberts^[Queensland University of Technology.]"
output:
  pdf_document:
    dev: jpeg
    fig_caption: yes
    number_sections: yes
    toc: no
    toc_depth: 3
geometry: margin = 1.2in
header-includes:
    - \usepackage{setspace}\doublespacing
link-citations: yes
csl: springer-vancouver.csl
fontsize: 11pt
bibliography: reviewbib.bib
yolo: yes
---

@@

\setcounter{page}{13}

@@

Cancer atlases communicate cancer statistics over geographic domains, typically with a choropleth map. A choropleth map subdivides these domains into administrative regions such as countries, states, or suburbs.  When communicating human-related statistics, the choropleth has a disadvantage in that it draws attention to sparsely populated rural areas to the neglect of small inner- city areas. The smaller geographic areas are important to consider if they are densely populated. Alternative map displays, such as a cartogram or a hexagon tile map, can shift the attention of map users from the large rural areas by decreasing their size on the map display. This means alternative displays can be more effective at accurately communicating spatial patterns across spatial areas. This study summarises current practices for cancer atlases and investigates the alternative map displays that could be used to accurately represent the distribution of cancer statistics across a population, as many cancer atlases lack appropriate displays for population statistics. It is recommended that alternative displays are included in cancer atlases. In addition, with the ease of today’s technology, user interaction with the displays is encouraged. Users should also be able to interactively display different statistics, such as incidence rate or relative incidence, or filter by demographic variables.

@@

<!-- Ways to visualize, whats our hypothesis -->
A choropleth map is the common display to visualize and communicate geospatial cancer statistics over geographic domain. Choropleth maps are a type of thematic map that show polygons for each of the groups of data points, representing the geographic units, where each polygon is shaded with a color according to the area-specific values of the statistic being conveyed. Visualizing this data is helpful as geographic patterns of disease may be obscured when reported in a table [@SAMGIS]. Providing a visual representation of cancer outcomes allows identification of geographic patterns of the disease that can then be addressed with public health policy and actions. The spatial distribution of the disease incidence can be examined using a choropleth and may reveal a trend in longitude or latitude, or rural vs urban, or coastal vs inland, or even specific hot spots of the disease. One of the key challenges with mapping spatial patterns of disease is the design of visualizations [@SE]. It is important to consider the strengths and weaknesses of designs, as visualizing diseases on maps is often the first step in exploratory spatial data analysis and helps in the formulation of hypotheses. This paper considers the current visualization techniques to communicate statistics to the public and their applications to cancer statistics. Alternative approaches are posed because they may be more effective than contemporary techniques. The limitations of the visualization methods, highlighting the differences and historic use of these displays is discussed.

<!-- Structure of the paper -->
The paper is structured as follows. The next section describes the choropleth map, which is the common approach to disease maps and presents examples of atlases in use today and discusses the limitations of the choropleth map. Section 3 describes alternative displays, including the cartogram, which is useful when the map has heterogeneously sized geographic units. Section 4 presents the limitations of the production and use of alternative displays. Disease maps are more useful when made interactive, and common options are described in Section 5, along with a discussion of benefits and disadvantages.

# Cancer atlases {#ch:public}
<!-- 360 words -->
<!-- Displaying of cancer data on choropleth maps -->
A cancer atlas is a map, or collection of maps, representing cancer incidence and mortality for a country, or group of countries. Atlases are key to developing hypotheses regarding areas with unusually high rates, and geographic correlations [@MACM]. The data collection methods across regions and the administrative control within regions lends itself to choropleth visualization. Cancer maps and atlases date back to Haviland's maps in 1875, and early work in US cancer atlases appearing in 1971 [@burbank]. The presentation of cancer statistics has increased with greater access to computational power and the availability of geographic information systems software [@SE].

<!-- Measures reported in cancer atlases -->
Cancer maps are effective tools for communicating incidence, survival, and mortality to a wide range of audiences, including the public and others not trained in statistical analysis. These visualizations enable non-expert audiences to interpret the outputs of sophisticated statistical analysis. Cruickshank (1947) as cited by S. D. Walter [@DMAHP], discusses using visuals as a 'formal statistical assessment of the spatial pattern'. Overwhelmingly, choropleth maps are visualisations chosen to communicate cancer statistics to members of the public and other non-expert audiences.

```{r choropleth-grid-create, message=FALSE, warning = FALSE, fig.cap = " A selection of choropleth cancer maps from online atlases that are publicly available. Maps of various countries were chosen: United Kingdom, Australia, Spain, USA, Canada, and display several different colour palettes and legends. These atllases are described in Table 1.", fig.align='center', fig.width=12, fig.height=10}
knitr::include_graphics("figures/choropleth_grid.png")
```

Table 1: <!--\label{tab:choropleths}--> A selection of choropleth cancer maps from online atlases.

|Fig.| Atlas | Statistic | Data source | 
|:---:|:------------|:----------------------|:----------------|  
|1a | The Environment and Health Atlas of England and Wales | relative risk for women developing lung cancer in England and Wales in 2010 [@EnvEnglandWales2]| [ Office for National Statistics (ONS) (England) and from the Welsh Cancer Intelligence and Surveillance Unit (WCISU) ](http://www.envhealthatlas.co.uk/homepage/)|
|1b | Globocan 2018: Estimated Cancer Incidence, Mortality and Prevalence Worldwide | age standardized incidence rates (per 100,000) for all invasive cancers for both men and women, aggregated at a national level for 2018 [@Globocan]| [World Health Organization's International Agency for Research on Cancer](https://www.iarc.fr/). |
|1c | Atlas of Cancer in Queensland | the relative incidence ratio of lung cancer in males in the state of QLD within Australia based on data from 1998 to 2007, Queensland Cancer Council[@QLDcancerAtlas] | [Queensland Cancer Registry](https://cancerqld.org.au/research/cancer-registries/queensland-cancer-registry/) |
|1d | Bowel Cancer Australia Atlas | the percentage of Australian males between 50 - 54 years of age diagnosed with bowel cancer in 2016 in Australia [@Bowel]. |  [Bowel Cancer Australia](https://www.bowelcanceraustralia.org/)|
|1e | United States Cancer Statistics: An Interactive Cancer Statistics Website | the incidence rate per 100,000, of all cancer types for men and women in the United States in 2016, aggregated at the state level [@USInteractive].|  *[Centers for Disease Control and Prevention](https://www.cdc.gov/)*, with data from state cancer registries. | 
|1f | Map of Cancer Mortality Rates in Spain | side by side maps of relative risk of lung cancer for men vs women for 2004 to 2008 [@cancerSpain].| [Map of cancer mortality rates in Spain](https://elpais.com/elpais/2014/10/06/media/1412612722_141933.html)|
|1g | Atlas of Childhood Cancer in Ontario | the incidence rate of childhood cancers per 100,000 (by census division) for children aged 0-14, in Ontario from 1995 to 2004 [@OntarioPediatric]. | [The Pediatric Oncology Group of Ontario Networked Information System](https://www.pogo.ca/research-data/pogonis-childhood-cancer-database/data-requests/)|

@@

Epidemiologists and statisticians have developed the statistics used to communicate the burden of cancer over several decades. Table 2<!--\ref{tab:measures}--> summarizes the measures commonly presented in published cancer atlases. Mortality rates are commonly presented as relative rates of risk across the population and age-adjusted to correct for the higher prevalence of cancers in older populations. As described in Howe [@HEDP], Englishman P. Stocks advanced the field of mortality statistics by introducing the standardized mortality ratios in the 1930s, which is an improvement on crude death rates.

Table 2: <!--\label{tab:measures}--> Common measures for reporting cancer information.

|Measure |Details| 
|:--------------------|:------------------------------------------------|  
|1. Count| Crude cancer counts|
|2. Rate per 100,000| Cancer incidence per 100,000 population|  
|3. IR (Incidence Ratio)| $(IR)_i=\frac{(Incidence\ Rate)_i}{Average\ Incidence\ Rate}$, |
|| The cancer incidence rate in region $i$ over the average cancer incidence rate for all of the regions|
|4. Age-Adjusted Rate per 100,000| Standardized by age structure or region|   
|5. Age-Adjusted Relative Risk| Standardized by age structure in each region $i$|  
|6. SIR (Standardized Incidence Ratio)| Incidence standardized by population at risk in each region $i$|
|7. Below or above Expected| An alternative expression of the SIR|
|8. RER | $RER = \frac{(Cancer\ related\ mortality)_i}{Average\ cancer\ related\ mortality}$|
|(Relative Excess Risk)| Represents the estimate of cancer-related mortality within five years of diagnosis. Also referred to as 'excess hazard ratio'|

<!-- 416 words -->
Roberts [@roberts2019communication] identified 33 cancer atlases published between 2010 and 2018. Each of these online atlases uses choropleth maps. All except one of these were published by non-commercial organizations, including not-for-profits, government, research organizations, advocacy groups or government-funded partnerships. Figure 1<!--\ref{fig:choropleth-grid-create}--> displays a subset of maps from these atlases, the selection varies in the geographies explored. Figure 1<!--\ref{fig:choropleth-grid-create}-->b shows Globocan 2018 [@Globocan] which explores Estimated Cancer Incidence, Mortality and Prevalence Worldwide using data sourced from cancer registries of each country. The Bowel Cancer Australia Atlas in Figure 1<!--\ref{fig:choropleth-grid-create}-->d presents an example of a cancer specific atlas -- it shows the average Standardized Incidence Ratio of colorectal cancer for Australian males from 2006 to 2010 [@Bowel]. Like many of the atlases examined, there is a choice of gender displayed in the Bowel Cancer Atlas. Gender is displayed in side-by-side maps in the Map of Cancer Mortality Rates in Spain (Figure 1f)<!--\ref{fig:choropleth-grid-create}--> [@cancerSpain].

Resolution of the maps varies greatly. Figure 1<!--\ref{fig:choropleth-grid-create}-->b shows global information at a national level. The United States Cancer Statistics [@USInteractive] shows data aggregated at the state level. The Environment and Health Atlas of England and Wales [@EnvEnglandWales2] (Figure 1<!--\ref{fig:choropleth-grid-create}-->a) shows the relative risk for women developing lung cancer  at a neighborhood (small-area) scale. The Atlas of Cancer in Queensland (Figure 1<!--\ref{fig:choropleth-grid-create}-->c) shows the relative incidence ratio of lung cancer in males for each Statistical Area at Level 2 [@abs2016] in the state of Queensland within Australia [@QLDcancerAtlas].

@@

<!-- psychology -->
Displaying familiar state boundaries can make a map easier to read [@CIBMUK] and allow viewers to infer the spatial relationships visually in the data using their mental model of the geography. The map users of disease displays may include researchers, the public, policymakers, and the media [@CPISACA]. For these users, the familiarity of the geography is a worthy consideration when presenting results of spatial analysis.

## Limitations of choropleth displays {#ch:chorolimit}

Australia presents an extreme case of an urban rural divide. The land mass occupied by urban electoral districts is only 10% of Australia, yet 90% of the population live in these urban areas [@ACTUC].

Choropleth maps provide a familiar display, which shows data in a geographically recognisable way. A disadvantage is that the different population and geographical sizes of administrative areas can attract attention to the shades of the underpopulated but large areas [@EI]. Skowronnek also [@BCM] discusses how choropleth maps suffer from area-size bias, as they give a 'stronger visual weight' to large  administrative units. The administrative boundaries used to define regions may limit a choropleth display, as this display unfaithfully represents the disease distribution across the region by obscuring small geographic areas. Sparsely populated rural areas are emphasized, whereas the areas representing inner city communities are very small. This is especially true for Australia.

>>>>

---
title: 'Mapping Cancer: the Potential of Cartograms and Alternative Map Displays'
author: Stephanie Kobakian^[Science and Engineering Faculty, Queensland University
  of Technology.], Dianne Cook^[Faculty of Econometrics and Business Statistics, Monash
  University.] and Jessie Roberts^[Queensland University of Technology.]
csl: springer-vancouver.csl
output:
  word_document:
    toc: no
    toc_depth: '3'
  pdf_document:
    dev: jpeg
    fig_caption: yes
    number_sections: yes
    toc: no
    toc_depth: 3
geometry: margin = 1.2in
header-includes: \usepackage{setspace}\doublespacing
link-citations: yes
fontsize: 11pt
bibliography: reviewbib.bib
yolo: yes
---

@@

@@

Cancer atlases communicate cancer statistics over geographic domains, typically with a choropleth map. A choropleth map subdivides these domains into administrative regions such as countries, states, or suburbs.  When communicating human-related statistics, the choropleth has a disadvantage in that it draws attention to sparsely populated rural areas to the neglect of small inner - city areas. The smaller geographic areas are important to consider if they are densely populated. Alternative map displays, such as a cartogram or a hexagon tile map, can shift the attention of map users from the large rural areas by decreasing their size on the map display. This means alternative displays can be more effective at accurately communicating spatial patterns across spatial areas. This study summarises current practices for cancer atlases and investigates the alternative map displays that could be used to accurately represent the distribution of cancer statistics across a population, as many cancer atlases lack appropriate displays for population statistics. It is recommended that alternative displays are included in cancer atlases. In addition, with the ease of today’s technology, user interaction with the displays is encouraged. Users should also be able to interactively display different statistics, such as incidence rate or relative incidence, or filter by demographic variables.

@@

<!-- Ways to visualize, whats our hypothesis -->
A choropleth map is the common display to visualize and communicate geospatial cancer statistics over geographic domain. Choropleth maps are a type of thematic map that show polygons for each of the groups of data points representing the geographic units, where each polygon is shaded with a color according to the area-specific values of the statistic being conveyed. Visualizing this data is helpful as geographic patterns of disease may be obscured when reported in a table [@SAMGIS]. Providing a visual representation of cancer outcomes allows identification of geographic patterns of the disease that can then be addressed with public health policy and actions. The spatial distribution of the disease incidence can be examined using a choropleth map display and may reveal a trend in longitude or latitude, or rural vs urban, or coastal vs inland, or even specific hot spots of the disease. One of the key challenges with mapping spatial patterns of disease is the design of visualizations [@SE]. It is important to consider the strengths and weaknesses of designs, as visualizing diseases on maps is often the first step in exploratory spatial data analysis and helps in the formulation of hypotheses.

<!-- # Cancer atlases -->
<!-- Displaying of cancer data on choropleth maps -->
A cancer atlas is a map, or collection of maps, representing cancer incidence and mortality for a country, or group of countries. Atlases are key to developing hypotheses regarding areas with unusually high rates, and geographic correlations [@MACM]. The data collection methods across regions and the administrative control within regions lends itself to choropleth visualization. Cancer maps and atlases date back to Haviland's maps in 1875, with more modern atlases directly evolving from early work in US cancer atlases, appearing in 1971 [@burbank]. The presentation of cancer statistics has increased with greater access to computational power and the availability of geographic information systems software [@SE].

<!-- Measures reported in cancer atlases -->
<!-- Structure of the paper -->

This paper considers the current visualization techniques to communicate statistics to the public and their applications to cancer statistics. Alternative approaches are posed because they may be more effective than contemporary techniques. The limitations of the visualization methods, highlighting the differences and historic use of these displays is discussed.

The paper is structured as follows. The next section introduces cancer atlases and the worldwide implementations. Choropleths are the common approach to disease maps and presents examples of atlases in use today and discusses the limitations of the choropleth map. Section 3 describes alternative displays, including the cartogram, which is useful when the map has heterogeneously sized geographic units. Section 4 presents the limitations of the production and use of alternative displays. Disease maps are more useful when made interactive, and common options are described in Section 5, along with a discussion of benefits and disadvantages.

@@

Epidemiologists and statisticians have developed the statistics used to communicate the burden of cancer over several decades. Table 1<!--\ref{tab:measures}--> summarizes the measures commonly presented in published cancer atlases. Mortality rates are commonly presented as relative rates of risk across the population and age-adjusted to correct for the higher prevalence of cancers in older populations. As described in Howe [@HEDP], Englishman P. Stocks advanced the field of mortality statistics by introducing the standardized mortality ratios in the 1930s, which is an improvement on crude death rates.

XXX
Some of the measures used to present cancer statistics are aggregations. The counts of cases in small areas can be difficult to obtain as they are often protected for privacy reasons. The information released and the statistics presented in worldwide cancer atlases are often modelled estimates. The measures described in table use Incidence as the statistic of interest, this is the amount of new cases and the same statistics can be used to describe the mortality, in this case Standardised Incidence Rate is analogous to the Standarised Mortality Rate.

<!--
Cancer mortality is the number of deaths from cancer in a specific population within a specific period of time, usually a year. It usually only includes deaths where cancer is mentioned as an underlying cause of death on death certificates.

Cancer mortality rates are a standard measure of the frequency of deaths within a specific period of time relative to a fixed population size, usually 100,000 people or expressed per million if rare.
-->

Table 1: <!--\label{tab:measures}--> Common measures for reporting cancer information.

|Measure |Details| Fig. |
|:--------------------|:------------------------------------------------|:--------------------|
|1. Count| Cancer incidence counts| |
|2. Incidence Rate (per 100,000) | The frequency of cases relative to a fixed population size, usually 100,000 people, within a specific period of time XXX [cite it](https://www.cancerresearchuk.org/health-professional/cancer-statistics/cancer-stats-explained/statistics-terminology-explained#heading-Eleven)|  |
|3. IR (Incidence Ratio)| The incidence rate in a region divided by the average  incidence rate for all of the regions| |
|4. Age-Standardized Rate per 100,000| Standardized by age structure or region|   |
|5. Age-Standardized Relative Risk| Standardized by age structure|  |
|6. SIR (Standardized Incidence Ratio)| The observed number of cases in a region divided by the expected number of cases, if that region had the same age-specific rates as a reference population.  |
|7. Excess Hazard Ratio | The estimate of cancer-related mortality within five years of diagnosis. |  |

<!-- psychology -->
Displaying familiar state boundaries can make a map easier to read [@CIBMUK] and allow viewers to infer the spatial relationships visually in the data using their mental model of the geography. The users of disease maps may include researchers, the public, policy makers, and the media [@CPISACA]. For these users, the familiarity of the geography is a worthy consideration when presenting results of spatial analysis.

Cancer maps are effective tools for communicating incidence, survival, and mortality to a wide range of audiences, including the public and others not trained in statistical analysis. These visualizations enable non-expert audiences to interpret the outputs of sophisticated statistical analysis. Cruickshank (1947) as cited by S. D. Walter [@DMAHP], discusses using visuals as a 'formal statistical assessment of the spatial pattern'. Overwhelmingly, choropleth maps are visualisations chosen to communicate cancer statistics to members of the public and other non-expert audiences.

```{r choropleth-grid-create, message=FALSE, warning = FALSE, fig.cap = " A selection of choropleth cancer maps from online atlases that are publicly available. Maps of various countries were chosen: United Kingdom, Australia, Spain, USA, Canada, and display several different colour palettes and legends. These atlases are described in Table 2.", fig.align='center', fig.width=12, fig.height=10}
knitr::include_graphics("figures/choropleth_grid.png")
```

Table 2: <!--\label{tab:choropleths}--> A selection of choropleth cancer maps from online atlases displayed in Figure 1.

| Fig. | Atlas | Data | Statistic | Source |
|------|---------------------------------------------------------------------------------------------------------|-----------------------------------------------------|------------------------------------------------------|---------------------------------------------------------------------------------------------------------------------------------------------------------------------------|
| a | The Environment and Health<br>Atlas of England and Wales<br>[@EnvEnglandWales2] | Sex: Women<br>Year: 2010<br>Cancer: Lung | Relative risk | [Office for National Statistics<br>(ONS) (England) and from the<br>Welsh Cancer Intelligence and<br>Surveillance Unit (WCISU)](http://www.envhealthatlas.co.uk/homepage/) |
| b | Globocan 2018: Estimated Cancer<br>Incidence, Mortality and <br><br>prevalence Worldwide<br>[@Globocan] | Sex: Both<br>Year: 2018<br>Cancer: All invasive | Age standardized<br>incidence rates<br>(per 100,000) | [World Health Organization's<br>International Agency for<br>Research on Cancer](https://www.iarc.fr/<br> |
| c | Atlas of Cancer in Queensland<br> [@QLDcancerAtlas] | Sex: Males<br>Year: 1998 - 2007<br>Cancer: Lung | Relative incidence<br>ratio | [Queensland Cancer Registry](https://cancerqld.org.au/research/cancer-registries/queensland-cancer-registry/) |
| d | Australian Cancer Atlas<br>[@ACA] | Sex: Males<br>Year: 2010 - 2014<br>Cancer: Lung | Average standardised<br>incidence rate | [Australian Cancer Atlas](https://atlas.cancer.org.au/) |
| e | United States Cancer<br>Statistics: An Interactive<br>Cancer Statistics Website<br> | Sex: Both<br>Year: 2016<br>Cancer: All types | Incidence rate per<br>100,000 | With data from state cancer<br>registries: <br>[Centers for Disease Control and Prevention](https://www.cdc.gov/) |
| f | Map of Cancer Mortality<br>Rates in Spain<br>[@cancerSpain] | Sex: Both<br>Year: 2004 - 2008<br>Cancer: All types | Relative risk | <br>[Map of cancer mortality rates<br>in Spain](https://elpais.com/elpais/2014/10/06/media/1412612722_141933.html |
| g | Atlas of Childhood<br>Cancer in Ontario<br>[@OntarioPediatric] | Sex: Both<br>Year: 1995 - 2004<br>Cancer: All types | Incidence rate per<br>100,00 | <br>[The Pediatric Oncology Group<br>of Ontario Networked Information<br>System](https://www.pogo.ca/research-data/pogonis-childhood-cancer-database/data-requests/) |

<!-- 416 words -->
A review of modern cancer atlases [@roberts2019communication] identified 33 cancer atlases published between 2010 and 2016. Atlases published between 2016 and 2018 have also been considered. Each of these online atlases uses choropleth maps. All except one of these were published by non-commercial organizations, including not-for-profits, government, research organizations, advocacy groups or government-funded partnerships. Figure 1<!--\ref{fig:choropleth-grid-create}--> displays a subset of maps from these atlases, the selection varies in the geographies explored. Figure 1<!--\ref{fig:choropleth-grid-create}-->b shows Globocan 2018 [@Globocan] which explores Estimated Cancer Incidence, Mortality and Prevalence Worldwide using estimates based on available country cancer registries. Many atlases allow users to view sex-specific distributions, however sex is displayed in side-by-side maps in the Map of Cancer Mortality Rates in Spain (Figure 1f)<!--\ref{fig:choropleth-grid-create}--> [@cancerSpain].

There is large variation in the resolution of the maps. Figure 1<!--\ref{fig:choropleth-grid-create}-->b shows global information at a national level. The United States Cancer Statistics [@USInteractive] shows data aggregated at the state level (n = 51). The Environment and Health Atlas of England and Wales [@EnvEnglandWales2] (Figure 1<!--\ref{fig:choropleth-grid-create}-->a) shows the relative risk for women developing lung cancer  at a neighborhood (small-area) scale (n = 8850). The Asutralian Cancer Atlas (Figure 1<!--\ref{fig:choropleth-grid-create}-->d) shows the relative incidence ratio of lung cancer in males for each Statistical Area at Level 2 [@abs2016] (n = 2292). The Atlas of Cancer in Queensland (Figure 1<!--\ref{fig:choropleth-grid-create}-->c) shows a subset of the Australian Statistical Areas at Level 2 [@abs2016] located in the state of Queensland within Australia [@QLDcancerAtlas] (n = 2292).

@@

Australia presents an extreme case of an urban rural divide. The land mass occupied by urban electoral districts is only 10% of Australia, yet 90% of the population live in these urban areas [@ACTUC]. Choropleth maps provide a familiar display, which shows data in a geographically recognisable way. A disadvantage is that the different population and geographical sizes of administrative areas can attract attention to the shades of the underpopulated but large areas [@EI]. Skowronnek also [@BCM] discusses how choropleth maps suffer from area-size bias, as they give a 'stronger visual weight' to large  administrative units. The administrative boundaries used to define regions may limit a choropleth display, as this display unfaithfully represents the disease distribution across the region by obscuring small geographic areas. Sparsely populated rural areas are emphasized, whereas the areas representing inner city communities are very small. This is especially true for Australia.

### 2020-05-14 submission2/submission2.Rmd
begin letter to editors using reviewer feedback
New figure 1 (include ACA)

hopeful

Table 1 and 2 switched, stats updated

**Reviewer suggestion:**
**Given the ethical and privacy considerations for small area data, it is not a straightforward process to obtain and report area-specific counts. I would suggest deleting this sentence and the previous one. **

We acknowledge that counts and incidence rates for small area data are priveledged information and have removed the sentences:

It is easy to provide counts or incidence rates of the diagnoses of these areas. This type of data is routinely collected for public health reasons and may be made available to the general public as a service to the community.

**Reviewer suggestion:**

**We would suggest that the aim of the manuscript could be phrased to highlight the current approaches and limitations of typical visualisation methods for cancer maps, along with outlining possible alternative methods and challenges in implementing these?**

We have restructed our original submission to provide a clear path from current approaches used, to the alternative approaches that should be explored. We highlight the limitations of using current display methods, to encourage the use of alternative display methods. We also present the challenges that may be faced when implementing the alternative designs.

The manuscript title has been changed:
"Cancer Applications of Choropleth Maps, and the Potential of Cartograms and Alternative Map Displays"

The manuscript now contains the following sections:

**Traditional approaches for cancer map displays:** This section discusses the choropleth map as the common map display used in cancer maps and atlases.
*Limitations of choropleth displays:* This section presents the possible issues with displaying cancer statistics on a choropleth map.
*Contemporary alternatives to choropleth maps:* We have shown that there are many alternative displays that highlight population dense regions.
*Challenges of alternative approaches:* Alternative displays also have limitations when displaying spatial distributions of cancer statistics.

We have inclued changes to reflect your minor notes:

*Editor suggestion:*
*The authors mention a focus on Atlases published between 2010 and 2015, but then include a bowel cancer atlas from 2016. *

*Also, within the bowel cancer atlas data, the reporting of percent of males aged 50-54 with bowel cancer in 2016 is incorrect; it should be the % of the ERP in each area that are males aged 50-54 years. *

We have changed the image used from the Bowel Cancer Atlas. To reflect this change the description of the statistic used in the bowel cancer atlas of Australia ahs been corrected:

Bowel Cancer Australia Atlas:
...shows the average Standardised Incidence Ratio of colorectal cancer for Australian males from 2006 to 2010 in Australia. It is published by Bowel Cancer Australia https://www.bowelcanceraustralia.org/.

The following section outlines comments made by the reviewer and contains our responses after the suggested changes were implemented.

*Reviewer A: *
*Major*
*Comment 1: *
The Abstract should be more focused- perhaps give an example of how cartogram mapping has been applied to data visualisation. What are the benefits of this approach and suggestions for the future.

*Reply 1:* 
The advantages of alternative mapping displays, including cartograms, has been included in the abstract. It clearly includes the changes to the map display made by alternative visualisation methods, and how they benefit the map user as they build an understanding of the spatial distribution.

*Changes in the text:*

Cancer atlases communicate cancer statistics over geographic domains, typically with a choropleth map. They subdivide these domains into administrative regions such as countries, states, or suburbs. When communicating human-related statistics, the choropleth has a disadvantage in that it draws attention to sparsely populated rural areas to the neglect of small inner city areas. The smaller geographic areas are important to consider if they are densely populated. Alternative map displays, such as a cartogram or a hexagon tile map, can shift the attention of map users from the large rural areas by decreasing their size on the map display. This means alternative displays can be more effective at accurately communicating spatial patterns across spatial areas, by highlighting areas of interest. It is recommended that alternative displays are included in cancer atlases. In addition, with the ease of today's technology, user interaction with the displays is encouraged. Users should also be able to interactively display different statistics, such as incidence rate or relative incidence.

*Comment 2: *

2. The key message seems to get lost in the Abstract. The Conclusion is vague.

*Reply 2:* 
Changes to the abstract have been made. Additionally, the conclusion has also been edited to be more direct in recommending the use or consideration of the alternatives presented.

*Changes in the text:* 
Please see the response to *Comment 1*.

*Comment 3: *

3. Entire paper needs substantial English language editing

*Reply 3:* 
Thank you for this point. Substantial time has been taken to improve the language used in this paper.

*Comment 4: *

4. Please avoid the use of sentences which have been used throughout the paper where you start with a reference number: such as “[4] discusses the use of (Map displays for disease data, Line 15-16) They do not make any sense and must be modified.

*Reply:* 
The referencing system was changed to implement the springer-vancouver style. We apologise for the issues this caused when exporting to a Word Document, the text has been modified appropriately.

*Comment 5: *

5. Please provide numbers for Figures and Tables

*Reply 5:* 
this begins the discussion, andNumbers for figures and tables have been corrected. This paper was produced using Rmarkdown, this introduced errors with numbering figures, sections and tables. All numbering has now been manually defined.

*Comment 6: *

6. Please clearly state the primary aim of this paper and structure it accordingly

*Reply 6:* 
The primary aim of this paper is to present alternative map displays as viable visualisation methods for presenting spatial distributions in cancer atlases. A choropleth map is commonly used in practice, it is described and it's limitations are highlighted. The alternative map displays address the limitations of the choropleth map. The methods and the limitations of these alternative displays are discussed.

*Comment 7: *

7. Conclusions must discuss the key focus and conclusions of this paper.

*Reply 7:* 
Small changes have been made to the conclusion that make the focus of the paper and recommendations clearer. The conclusion has three parts (1) mapping practices, and recommendations, (2) statistics applied and the pros and cons, and (3) user interaction with online maps and the benefits of incorporating this.

*Minor Comments*

- Please check the use of past and present tense
- There are many minor grammatical errors that need to checked.
- Please edit the last section of the Introduction. Either number each section and then refer to them or find another way to appropriately refer to each section.

>>>>

**Reviewer suggestion:**
**Could we structure paper as:**
**Introduction (also introduces cancer atlases)**
**Measures mapped (info around Table 2)**
**Visualisation approaches**
**- Choropleth (mention advant and disadvant, along with figures from cancer atlases)**
**- Cartograms**
**- Tile maps**
**- Geofacets**
**- Multivariate maps**
**Comparison & critique of alternative methods**
**Additional considerations**
**- Associated graphs**
**- Displaying uncertainty measures**
**- Interactivity**
**Recommendations for mapping cancer**
**Conclusion**

We have rearranged the sections from the "Introduction" to "Traditional approaches for cancer map displays" in this submission at the request of the reviewer. The discussion of choropleth maps considers the advantages and disadvantages now occurs in the same section.

**Reviewer suggestion:**
**Suggest also noting development of statistical methods and software**

**Reviewer suggestion:**
**Give the actual reference for Cruikshank 1947**

This quote has been removed.

**Reviewer suggestion:**
**Suggest including Australian Cancer Atlas**
We have replaced the Bowel Cancer Australia Atlas with the Australian Cancer Atlas.

**Reviewer suggestion:**
**This table might benefit from an additional column, which gives an example of a cancer map using this measure. Also, some of these measures can be calculated.**

Several suggestions were made regarding Table 2: Common measures for reporting cancer information.
The additional information requested is communicated more effectively using a paragraph style format.
